---
title: Dietary fiber content in common bean from the middle
  american gene pool
author: "skammlade"
date: "17 June 2016"
output:
  pdf_document:
    fig_caption: yes
    includes:
      in_header: preamble_fig_env.tex
    latex_engine: xelatex
    toc: yes
    toc_depth: 2
  html_document:
    toc: yes
  word_document:
    fig_caption: yes
    toc: yes
header-includes:
- \usepackage{booktabs}
- \usepackage{longtable}
- \captionsetup[table]{singlelinecheck=false,justification=raggedright,position=top,skip=50pt}
graphics: yes
sansfont: Calibri Light
fontsize: 12pt
---

\renewcommand{\sfdefault}{phv}

## Introduction

```{r echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
fiber<- read.csv("fiber.csv")
#subset fiber data to remove black mottle, flor de mayo, red mottle, carioca, and tan
fibersub <- subset(fiber, mktclass=='GN' | mktclass=='pink' | mktclass=='pinto' | mktclass=='small red' | mktclass=='black' | mktclass=='small white')
```

`r nrow(fibersub)/2` entries from the BeanCAP study were grown in Fort Collins, CO in 2015. All entries originated from the Middle American gene pool and were further subdivided into three races and twelve market classes (Table 1). Beans were analyzed for various fiber components to determine the effect of marketclass and race (Table 2).

| Race | Market Class | No. Entries | 
| ------------ | ----------- | ------------ | 
| Durango | | **`r sum(fibersub$race=="durango", na.rm=TRUE)/2`** |
| | GN | `r sum(fibersub$mktclass=="GN", na.rm=TRUE)/2` |
| | pink | `r sum(fibersub$mktclass=="pink", na.rm=TRUE)/2` |
| | pinto | `r sum(fibersub$mktclass=="pinto", na.rm=TRUE)/2` |
| | small red | `r sum(fibersub$mktclass=="small red", na.rm=TRUE)/2` | 
| Mesoamerican | | **`r sum(fibersub$race=="mesoamerican", na.rm=TRUE)/2`** |
| | black |   `r sum(fibersub$mktclass=="black", na.rm=TRUE)/2` |
| | small white     |  `r sum(fibersub$mktclass=="small white", na.rm=TRUE)/2` |

Table: Number of entries by race and market class included in the study.

Fiber component | Description
--------------------|-------------------------------------------
Soluble | Dissolves in water
Insoluble | Does not dissolve in water
Raffinose | A trisaccharide
Stachyose | A tetrasaccharide
Verbascose | A pentasachharide
Total Oligosaccharides | Sum of verbascose, raffinose, and stachyose
Total Dietary Fiber | Sum of soluble fiber, insoluble fiber, and total oligosaccharides

Table: Fiber components measured in the study. Faffinose, stachyose, and verbascose are grouped together broadly as total oligosaccharides. These carbohydrates are not digested in the small intestine and are instead passed into the large intestine to be fermented by microflora with the production of gas.

The red mottle market class had the greatest mean total dietary fiber, as well as the greatest mean verbascose, stachyose, total oligosaccharide, and insoluble fiber. The black mottle market class had the greatest mean raffinose and soluble dietary fiber content. Both of these market classes included only one entry each ('Orcha' a black mottle; 'Ind. Jamaica Red' a red mottle) making these findings not as insightful. A one-way analysis of variance revealed a significant effect (p<0.5) of market class on all fiber components.

A one-way analysis of variance revealed a significant effect (p<0.5) of race on all fiber components except total oligosachharide content. Durango and jalisco races had the greatest mean verbascose and stachyose content, mesoamerican race had the greatest mean raffinose content. Jalisco had the greatest mean insoluble DF, soluble DF, and total dietary fiber content.

## Results: Entry Rankings

Tables below show the five entries with the greatest and five entries with the lowest values of each fiber component averaged across the two replications including standard deviation (sd), standard error (se) and confindence intervals (ci).

## -- Insoluble Dietary Fiber (IDF) 
```{r echo=FALSE, message=FALSE, warning=FALSE}
library(data.table)
library(knitr)

fiber<- read.csv("fiber.csv")
#subset fiber data to remove black mottle, flor de mayo, red mottle, carioca, and tan
fibersub <- subset(fiber, mktclass=='GN' | mktclass=='pink' | mktclass=='pinto' | mktclass=='small red' | mktclass=='black' | mktclass=='small white')

source('summarySE.R')
summ.IDF <-summarySE(fibersub, measurevar="IDF", groupvars=c("Entry","mktclass","race"))
summ.IDF <- data.frame(summ.IDF)
#summ.f <- summ.F[fiber$ftype=="IDF", ]
f.IDF <- subset(summ.IDF, select=c(Entry, mktclass, race, IDF, sd, se, ci))
newdata <- f.IDF[order(-f.IDF$IDF), ]
newdata[,4:7] <- round(newdata[,4:7], 2)
DT <- data.table(newdata)
head <- head(DT, 5)
kable(head, caption="Five entries with greatest IDF content.")
tail <- tail(DT,5)
kable(tail, caption="Five entries with lowest IDF content.")
```

\pagebreak

## --  Soluble Dietary Fiber (SDF)
```{r echo=FALSE, message=FALSE, warning=FALSE}
library(data.table)
library(knitr)

fiber<- read.csv("fiber.csv")
#subset fiber data to remove black mottle, flor de mayo, red mottle, carioca, and tan
fibersub <- subset(fiber, mktclass=='GN' | mktclass=='pink' | mktclass=='pinto' | mktclass=='small red' | mktclass=='black' | mktclass=='small white')

source('summarySE.R')
summ.SDF <-summarySE(fibersub, measurevar="SDF", groupvars=c("Entry","mktclass","race"))
summ.SDF <- data.frame(summ.SDF)
#summ.f <- summ.F[fiber$ftype=="SDF", ]
f.SDF <- subset(summ.SDF, select=c(Entry, mktclass, race, SDF, sd, se, ci))
newdata <- f.SDF[order(-f.SDF$SDF), ]
newdata[,4:7] <- round(newdata[,4:7], 2)
DT <- data.table(newdata)
head <- head(DT, 5)
kable(head, caption="Five entries with greatest SDF content.")
tail <- tail(DT,5)
kable(tail, caption="Five entries with lowest SDF content.")
```

## --  Raffinose (Raff)
```{r echo=FALSE, message=FALSE, warning=FALSE}
library(data.table)
library(knitr)

fiber<- read.csv("fiber.csv")
#subset fiber data to remove black mottle, flor de mayo, red mottle, carioca, and tan
fibersub <- subset(fiber, mktclass=='GN' | mktclass=='pink' | mktclass=='pinto' | mktclass=='small red' | mktclass=='black' | mktclass=='small white')

source('summarySE.R')
summ.Raff <-summarySE(fibersub, measurevar="Raff", groupvars=c("Entry","mktclass","race"))
summ.Raff <- data.frame(summ.Raff)
#summ.f <- summ.F[fiber$ftype=="Raff", ]
f.Raff <- subset(summ.Raff, select=c(Entry, mktclass, race, Raff, sd, se, ci))
newdata <- f.Raff[order(-f.Raff$Raff), ]
newdata[,4:7] <- round(newdata[,4:7], 2)
DT <- data.table(newdata)
head <- head(DT, 5)
kable(head, caption="Five entries with greatest raffinose content.")
tail <- tail(DT,5)
kable(tail, caption="Five entries with lowest raffinose content.")
```

## --  Stachyose (Stach)
```{r echo=FALSE, message=FALSE, warning=FALSE}
library(data.table)
library(knitr)

fiber<- read.csv("fiber.csv")
#subset fiber data to remove black mottle, flor de mayo, red mottle, carioca, and tan
fibersub <- subset(fiber, mktclass=='GN' | mktclass=='pink' | mktclass=='pinto' | mktclass=='small red' | mktclass=='black' | mktclass=='small white')

source('summarySE.R')
summ.stach <-summarySE(fibersub, measurevar="Stach", groupvars=c("Entry","mktclass","race"))
summ.stach <- data.frame(summ.stach)
#summ.f <- summ.F[fiber$ftype=="Stach", ]
f.stach <- subset(summ.stach, select=c(Entry, mktclass, race, Stach, sd, se, ci))
newdata <- f.stach[order(-f.stach$Stach), ]
newdata[,4:7] <- round(newdata[,4:7], 2)
DT <- data.table(newdata)
head <- head(DT, 5)
kable(head, caption="Five entries with greatest stachyose content.")
tail <- tail(DT,5)
kable(tail, caption="Five entries with lowest stachyose content.")
```

## --  Verbascose (Verb)
```{r echo=FALSE, message=FALSE, warning=FALSE}
library(data.table)
library(knitr)

fiber<- read.csv("fiber.csv")
#subset fiber data to remove black mottle, flor de mayo, red mottle, carioca, and tan
fibersub <- subset(fiber, mktclass=='GN' | mktclass=='pink' | mktclass=='pinto' | mktclass=='small red' | mktclass=='black' | mktclass=='small white')

source('summarySE.R')
summ.Verb <-summarySE(fibersub, measurevar="Verb", groupvars=c("Entry","mktclass","race"))
summ.Verb <- data.frame(summ.Verb)
#summ.f <- summ.F[fiber$ftype=="Verb", ]
f.Verb <- subset(summ.Verb, select=c(Entry, mktclass, race, Verb, sd, se, ci))
newdata <- f.Verb[order(-f.Verb$Verb), ]
newdata[,4:7] <- round(newdata[,4:7], 3)
DT <- data.table(newdata)
head <- head(DT, 5)
kable(head, caption="Five entries with greatest verbascose content.")
tail <- tail(DT,5)
kable(tail, caption="Five entries with lowest verbascose content.")
```

## --  Total Oligosachharides (TOligos)
```{r echo=FALSE, message=FALSE, warning=FALSE}
library(data.table)
library(knitr)

fiber<- read.csv("fiber.csv")
#subset fiber data to remove black mottle, flor de mayo, red mottle, carioca, and tan
fibersub <- subset(fiber, mktclass=='GN' | mktclass=='pink' | mktclass=='pinto' | mktclass=='small red' | mktclass=='black' | mktclass=='small white')

source('summarySE.R')
summ.TOligos <-summarySE(fibersub, measurevar="TOligos", groupvars=c("Entry","mktclass","race"))
summ.TOligos <- data.frame(summ.TOligos)
#summ.f <- summ.F[fiber$ftype=="TOligos", ]
f.TOligos <- subset(summ.TOligos, select=c(Entry, mktclass, race, TOligos, sd, se, ci))
newdata <- f.TOligos[order(-f.TOligos$TOligos), ]
newdata[,4:7] <- round(newdata[,4:7], 2)
DT <- data.table(newdata)
head <- head(DT, 5)
kable(head, caption="Five entries with greatest oligosaccharide content.")
tail <- tail(DT,5)
kable(tail, caption="Five entries with lowest oligosaccharide content.")
```

## --  Total Dietary Fiber (TDF) 

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Table #. Total Dietary Fiber top 5 and bottom 5."}
library(data.table)
library(knitr)

fiber<- read.csv("fiber.csv")
#subset fiber data to remove black mottle, flor de mayo, red mottle, carioca, and tan
fibersub <- subset(fiber, mktclass=='GN' | mktclass=='pink' | mktclass=='pinto' | mktclass=='small red' | mktclass=='black' | mktclass=='small white')

source('summarySE.R')
summ.TDF <- summarySE(fibersub, measurevar="TDF", groupvars=c("Entry","mktclass","race"))
summ.TDF <- data.frame(summ.TDF)
#summ.f <- summ.F[fiber$ftype=="TDF", ]
f.TDF <- subset(summ.TDF, select=c(Entry, mktclass, race, TDF, sd, se, ci))
newdata <- f.TDF[order(-f.TDF$TDF), ]
newdata[,4:7] <- round(newdata[,4:7], 2)
DT <- data.table(newdata)
head <- head(DT, 5)
kable(head, caption="Five entries with greatest total DF content.")
tail <- tail(DT,5)
kable(tail, caption="Five entries with lowest total DF content.")
```


## Results: Marketclass

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(data.table)
fiber.one<- read.csv("fiber_facet.csv")
```


##   --   Boxplots

```{r echo=FALSE, message=FALSE, warning=FALSE}
source('subset.race.R')
```

```{r fig.width = 8, fig.height=8, fig.align='center', echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Oligosaccharide content in 282 common bean genotypes by marketclass (A) Raffinose, (B) Stachyose, (C) Verbascose, and (D) total oligosaccharides (= Raff + Stach + Verb)."}
##Boxplots by Marketclass: Oligosaccharides
source('boxplot.mktclass.R')
plot_grid( p.Raff.mkt, p.Stach.mkt, p.Verb.mkt, p.TOligos.mkt, labels = c("A", "B", "C", "D"), ncol = 2)
```

\pagebreak

| Marketclass | Verbascose | Raffinose | Stachyose | Total Oligosaccharides |
---------|--------|--------|---------|---------|
black | `r round(mean(verb.black$value),3)` a | `r round(mean(raff.black$value),3)` c | `r round(mean(stach.black$value),2)` b | `r round(mean(toligos.black$value),2)` bc |
black mottle | `r round(mean(verb.blackmottle$value),3)` ab | **`r round(mean(raff.blackmottle$value),3)` bc** | `r round(mean(stach.blackmottle$value),2)` ab | `r round(mean(toligos.blackmottle$value),2)` ab |
carioca | `r round(mean(verb.carioca$value),3)` ab | `r round(mean(raff.carioca$value),3)` bc | `r round(mean(stach.carioca$value),2)` ab | `r round(mean(toligos.carioca$value),2)` abc |
flor de mayo | `r round(mean(verb.flordemayo$value),3)` abd | `r round(mean(raff.flordemayo$value),3)` bc | `r round(mean(stach.flordemayo$value),2)` ab | `r round(mean(toligos.flordemayo$value),2)` abc |
GN | `r round(mean(verb.GN$value),3)` b | `r round(mean(raff.GN$value),3)` ac | `r round(mean(stach.GN$value),2)` b | `r round(mean(toligos.GN$value),2)` bc |
navy | `r round(mean(verb.navy$value),3)` ac | `r round(mean(raff.navy$value),3)` ac | `r round(mean(stach.navy$value),2)` a | `r round(mean(toligos.navy$value),2)` a |
pink | `r round(mean(verb.pink$value),3)` bc | `r round(mean(raff.pink$value),3)` b | `r round(mean(stach.pink$value),2)` ab | `r round(mean(toligos.pink$value),2)` ab |
pinto | `r round(mean(verb.pinto$value),3)` b | `r round(mean(raff.pinto$value),3)` b | `r round(mean(stach.pinto$value),2)` b | `r round(mean(toligos.pinto$value),2)` b |
red mottle | **`r round(mean(verb.redmottle$value),3)` d** | `r round(mean(raff.redmottle$value),3)` ab | **`r round(mean(stach.redmottle$value),2)` c** | **`r round(mean(toligos.redmottle$value),2)` c** |
small red | `r round(mean(verb.smallred$value),3)` b | `r round(mean(raff.smallred$value),3)` ab | `r round(mean(stach.smallred$value),2)` b | `r round(mean(toligos.smallred$value),2)` bc |
small white | `r round(mean(verb.smallwhite$value),3)` ab | `r round(mean(raff.smallwhite$value),3)` bc | `r round(mean(stach.smallwhite$value),2)` ab | `r round(mean(toligos.smallwhite$value),2)` abc |
tan | `r round(mean(verb.tan$value),3)` ab | `r round(mean(raff.tan$value),3)` bc | `r round(mean(stach.tan$value),2)` ab | `r round(mean(toligos.tan$value),2)` ab |

Table: Mean fiber component content by market class. Values with different letters differ significantly between market classes (p<0.05). Bolded values are the greatest among market classes.

\pagebreak

```{r fig.width = 8, fig.height=5, fig.align='center', echo=FALSE, message=FALSE, warning=FALSE, fig.cap="(A) Soluble and (B) insoluble dietary fiber in 282 common bean genotypes by marketclass."}
##Boxplots by Marketclass: Soluble and Insoluble DF
source('boxplot.mktclass.R')
plot_grid(p.IDF.mkt, p.SDF.mkt, labels = c("A", "B"), ncol = 2)
```

| Marketclass | Insoluble DF | Soluble DF
---------|--------|--------
black | `r round(mean(idf.black$value),2)` cd | `r round(mean(sdf.black$value),2)` b 
black mottle | `r round(mean(idf.blackmottle$value),2)` abc | **`r round(mean(sdf.blackmottle$value),2)` ab** 
carioca | `r round(mean(idf.carioca$value),2)` abc | `r round(mean(sdf.carioca$value),2)` ab
flor de mayo | `r round(mean(idf.flordemayo$value),2)` abc | `r round(mean(sdf.flordemayo$value),2)` ab
GN | `r round(mean(idf.GN$value),2)` ab | `r round(mean(sdf.GN$value),2)` ab 
navy | `r round(mean(idf.navy$value),2)` a | `r round(mean(sdf.navy$value),2)` ab 
pink | `r round(mean(idf.pink$value),2)` abc | `r round(mean(sdf.pink$value),2)` a 
pinto | `r round(mean(idf.pinto$value),2)` bd | `r round(mean(sdf.pinto$value),2)` b 
red mottle | **`r round(mean(idf.redmottle$value),2)` abc** | `r round(mean(sdf.redmottle$value),2)` ab 
small red | `r round(mean(idf.smallred$value),2)` c | `r round(mean(sdf.smallred$value),2)` ab 
small white | `r round(mean(idf.smallwhite$value),2)` abc | `r round(mean(sdf.smallwhite$value),2)` ab 
tan | `r round(mean(idf.tan$value),2)` abc | `r round(mean(sdf.tan$value),2)` ab 

Table: Mean fiber component content by market class. Values with different letters differ significantly between market classes (p<0.05). Bolded values are the greatest among market classes.

```{r fig.width = 4, fig.height=4.5, fig.align='center', echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Total Dietary Fiber content in 282 common bean genotypes by marketclass."}
##Boxplots by Marketclass: Total Dietary Fiber
source('boxplot.mktclass.R')
p.TDF.mkt
```

| Marketclass | Total DF 
---------|--------
black | `r round(mean(tdf.black$value),2)` bc 
black mottle | `r round(mean(tdf.blackmottle$value),2)` abc 
carioca | `r round(mean(tdf.carioca$value),2)` abc 
flor de mayo | `r round(mean(tdf.flordemayo$value),2)` abc
GN | `r round(mean(tdf.GN$value),2)` cd 
navy | `r round(mean(tdf.navy$value),2)` a 
pink | `r round(mean(tdf.pink$value),2)` bd 
pinto | `r round(mean(tdf.pinto$value),2)` ac 
red mottle | **`r round(mean(tdf.redmottle$value),2)` bc**
small red | `r round(mean(tdf.smallred$value),2)` b 
small white | `r round(mean(tdf.smallwhite$value),2)` abc 
tan | `r round(mean(tdf.tan$value),2)` abc 

Table: Mean total dietary fiber component content by market class. Values with different letters differ significantly between market classes (p<0.05).

## Results: Race

```{r echo=FALSE, message=FALSE, warning=FALSE}
source('subset.race_real.R')
```

##   --   Boxplots

```{r fig.width = 8, fig.height=6.5, fig.align='center', echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Oligosaccharide content in 282 common bean genotypes by race (A) Raffinose, (B) Stachyose, (C) Verbascose, and (D) total oligosaccharides (= Raff + Stach + Verb)."}
##Boxplots by race: Oligosaccharides
source('boxplot.race.R')
plot_grid(p.Raff.race, p.Stach.race, p.Verb.race, p.TOligos.race, labels = c("A", "B", "C", "D"), ncol = 2, align="w")
```

| Race | Verbascose | Raffinose | Stachyose | Total Oligosaccharides |
---------|--------:|--------:|---------:|---------:|
Durango | `r round(mean(verb.durango$value),3)` b | `r round(mean(raff.durango$value),3)` a | `r round(mean(stach.durango$value),2)` b | `r round(mean(toligos.durango$value),2)` a |
Jalisco | `r round(mean(verb.jalisco$value),3)` b | `r round(mean(raff.jalisco$value),3)` a | `r round(mean(stach.jalisco$value),2)` ab | `r round(mean(toligos.jalisco$value),2)` a |
Mesoamerican | `r round(mean(verb.meso$value),3)` a | `r round(mean(raff.meso$value),3)` b | `r round(mean(stach.meso$value),2)` a | `r round(mean(toligos.meso$value),2)` a |

Table: Mean fiber component content by race. Values with different letters differ significantly between races (p<0.05).

```{r fig.width = 8, fig.height=4, fig.align='center', echo=FALSE, message=FALSE, warning=FALSE, fig.cap="(A) Soluble and (B) insoluble dietary fiber in 282 common bean genotypes by race,"}
##Boxplots by race: Soluble and Insoluble DF
source('boxplot.race.R')
plot_grid(p.IDF.race, p.SDF.race, labels = c("A", "B"), ncol = 2, align="w")
```

| Race | Insoluble DF | Soluble DF | 
---------|--------:|--------:|
Durango | `r round(mean(idf.durango$value),2)` a | `r round(mean(sdf.durango$value),2)` a | 
Jalisco | `r round(mean(idf.jalisco$value),2)` b | `r round(mean(sdf.jalisco$value),2)` b | 
Mesoamerican | `r round(mean(idf.meso$value),2)` a | `r round(mean(sdf.meso$value),2)` a |

Table: Mean fiber component content by race. Values with different letters differ significantly between races (p<0.05).

```{r fig.width = 4, fig.height=4, fig.align='center', echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Total Dietary Fiber content in 282 common bean genotypes by race"}
##Boxplots by race: Total Dietary Fiber
source('boxplot.race.R')
p.TDF.race
```

| Race | Total DF |
---------|--------:|
Durango | `r round(mean(tdf.durango$value),2)` a | 
Jalisco | `r round(mean(tdf.jalisco$value),2)` b | 
Mesoamerican | `r round(mean(tdf.meso$value),2)` a |

Table: Mean total dietary fiber content by race. Values with different letters differ significantly between races (p<0.05).

\pagebreak

## Statistical Methods

Statistical analyses, figures, tables, and reporting was completed using RStudio Version 0.98.1062.

## --  *ANOVA Assumptions*

Insoluble fiber, soluble fiber, raffinose, stachyose, and verbascose data were concluded to meet the assumptions of normality of residuals and equality of variance after plotting qqplots, and residuals versus predicted values. Diagnostic plots are shown below.

```{r fig.width = 8, fig.height=10, echo=FALSE, message=FALSE, warning=FALSE, include=FALSE}
source('plot.test.R')
```

```{r fig.width = 8, fig.height=6, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Residual versus fitted and q-q plots of insoluble and soluble DF. The insoluble DF plots look to reveal non-homogeneity of variance, but both the log and square root transformations did not improve it enough to warrant transforming the data."}
library(cowplot)
plot_grid(p1,p2,p3,p4, ncol=2, labels = c("A", "B", "C", "D"))
```

```{r fig.width = 8, fig.height=9, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Residual versus fitted and q-q plots of verbascose, raffinose, and stachyose. The raffinose plots look to reveal non-homogeneity of variance, but both the log and square root transformations did not improve it enough to warrant transforming the data."}
library(cowplot)
plot_grid(p5,p6,p7,p8,p9,p10, ncol=2, labels = c("A", "B", "C", "D", "E", "F"))
```

## -- *ANOVA*

One-way analysis of variance was conducted to determine the effect of 1) rep, 2) race, and 3) marketclass. Rep was treated as random, race and marketclass were treated as fixed with marketclass nested within race.

#### *1) Market class*

Differences between market classes were significant (p<0.05) across all fiber components.

```{r echo=FALSE, message=FALSE, warning=FALSE}
source('anova.nest.R')
library(xtable, knitr)
kable(xtable(OneWayFit.IDF), caption="ANOVA table testing insoluble DF.", digits=4) 
kable(xtable(OneWayFit.SDF), caption="ANOVA table testing soluble DF.", digits=4)
kable(xtable(OneWayFit.Raff), caption="ANOVA table testing raffinose.", digits=4)
kable(xtable(OneWayFit.Stach), caption="ANOVA table testing stachyose.", digits=4)
kable(xtable(OneWayFit.Verb), caption="ANOVA table testing verbascose.", digits=4)
kable(xtable(OneWayFit.TOligos), caption="ANOVA table testing total oligosaccharides.", digits=4)
kable(xtable(OneWayFit.TDF), caption="ANOVA table testing total DF.", digits=4)
```


## -- *Pairwise Comparisons*

Tukey adjusted pairwise comparisons were made between each market class and each race for each fiber component using the TukeyHSD funcion. These results were used in mean summary tables below boxplots.

```{r echo=TRUE, message=FALSE, warning=FALSE, results='hide'}
OneWayFit <- aov(IDF~race, data=fiber)
TukeyHSD(OneWayFit)

library(multcomp)
PairComps <- glht(OneWayFit, linfct=mcp(race="Tukey"))
summary(PairComps)
cld(PairComps)
```