---
title: Dietary fiber content in common bean from the middle
  american gene pool
author: "skammlade"
date: "17 June 2016"
output:
  pdf_document:
    fig_caption: yes
    includes:
      in_header: preamble_fig_env.tex
    latex_engine: xelatex
    toc: yes
    toc_depth: 2
  html_document:
    toc: yes
  word_document:
    fig_caption: yes
    toc: yes
header-includes:
- \usepackage{booktabs}
- \usepackage{longtable}
- \captionsetup[table]{singlelinecheck=false,justification=raggedright,position=top,skip=50pt}
graphics: yes
sansfont: Calibri Light
fontsize: 12pt
---

\renewcommand{\sfdefault}{phv}

## Introduction

```{r echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
fiber<- read.csv("fiber.csv")
#subset fiber data to remove black mottle, flor de mayo, red mottle, carioca, and tan
fibersub <- subset(fiber, mktclass=='GN' | mktclass=='pink' | mktclass=='pinto' | mktclass=='small red' | mktclass=='black' | mktclass=='small white')
```

`r nrow(fibersub)/2` entries from the BeanCAP study were grown in Fort Collins, CO in 2015. All entries originated from the Middle American gene pool and were further subdivided into two races and six market classes after removing lesser important market classes (Table 1). Beans were analyzed for various fiber components to determine the effect of marketclass and race (Table 2).

| Race | Market Class | No. Entries | 
| ------------ | ----------- | ------------ | 
| Durango | | **`r sum(fibersub$race=="durango", na.rm=TRUE)/2`** |
| | GN | `r sum(fibersub$mktclass=="GN", na.rm=TRUE)/2` |
| | pink | `r sum(fibersub$mktclass=="pink", na.rm=TRUE)/2` |
| | pinto | `r sum(fibersub$mktclass=="pinto", na.rm=TRUE)/2` |
| | small red | `r sum(fibersub$mktclass=="small red", na.rm=TRUE)/2` | 
| Mesoamerican | | **`r sum(fibersub$race=="mesoamerican", na.rm=TRUE)/2`** |
| | black |   `r sum(fibersub$mktclass=="black", na.rm=TRUE)/2` |
| | small white     |  `r sum(fibersub$mktclass=="small white", na.rm=TRUE)/2` |

Table: Number of entries by race and market class included in the study.

Fiber component | Description
--------------------|-------------------------------------------
Soluble | Dissolves in water
Insoluble | Does not dissolve in water
Raffinose | A trisaccharide
Stachyose | A tetrasaccharide
Verbascose | A pentasachharide
Total Oligosaccharides | Sum of verbascose, raffinose, and stachyose
Total Dietary Fiber | Sum of soluble fiber, insoluble fiber, and total oligosaccharides

Table: Fiber components measured in the study. Faffinose, stachyose, and verbascose are grouped together broadly as total oligosaccharides. These carbohydrates are not digested in the small intestine and are instead passed into the large intestine to be fermented by microflora with the production of gas.

```{r echo=FALSE, message=FALSE, warning=FALSE}
source('summarySE.R')
source('ranking.R')
library(knitr)
```

## Results: Soluble (SDF) and Insoluble (IDF) Dietary Fiber

## -- Entry Rankings 
```{r echo=FALSE, message=FALSE, warning=FALSE}
kable(data.table(rank.SDF, rank.IDF), caption="Five entries with largest and smallest IDF and SDF content.")
```

## Results: Raffinose Pathway of Oligosaccharides

## -- Entry Rankings
```{r echo=FALSE, message=FALSE, warning=FALSE}
kable(data.table(rank.Raff, rank.Stach, rank.Verb, rank.TOligos), caption="Five entries with greatest oligosaccharide content from raffinose pathway.")
```

## -- Oligosachharide mktclass and race means

```{r fig.width = 8, fig.height=8, fig.align='center', echo=FALSE, message=FALSE, warning=FALSE, fig.cap="(A) Boxplot of Entry means by origin and (B) lineplot of entry means by origin showing gxe."}
source('barplot.raffpath.R')
source('lineplot.oligo.race.R')
plot_grid(p.raffpath, p.oli.race, ncol=1, labels=c("A", "B"))
```




## Results: Total Dietary Fiber

## --  Entry Rankings
```{r echo=FALSE, message=FALSE, warning=FALSE}
kable(rank.TDF, caption="Five entries with greatest and lowest TDF content.")
```

## Results: GxE

## --  Soluble (SDF) and Insoluble (IDF) Fiber

```{r fig.width = 8, fig.height=6, fig.align='center', echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Entry means by origin and lineplot of entry means by origin showing gxe of (A) Soluble and (B) insoluble fiber. Bars with the same letter do not differe significantly between locations. An analysis of variance revealed there is not a significant effect of location on either soluble (p=0.8717) or insoluble fiber (p=0.1471). There is a significant effect of gxe on both soluble fiber and insoble fiber (p=0.0029 and p=0.0003 respectively)."}
source('lineplot.gxe.R')
source('barplot.gxe.origin.R')
plot_grid(pbar.SDF.origin, p.gxe.SDF, pbar.IDF.origin, p.gxe.IDF, labels = c("A", "", "B", ""), ncol = 2, align="hv")
```

## --  Raffinose Oligosaccharide Pathway

```{r fig.width = 8, fig.height=8, fig.align='center', echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Entry means by origin and lineplot of entry means by origin showing gxe of (A) raffinose, (B) stachyose, (C) verbascose, and (D) total oligosaccharides. Bars with the same letter do not differe significantly between locations. An analysis of variance revealed there is not a significant effect of location nor gxe on either components."}
source('lineplot.gxe.R')
source('barplot.gxe.origin.R')
plot_grid(pbar.Raff.origin, p.gxe.Raff, pbar.Stach.origin, p.gxe.Stach, pbar.Verb.origin, p.gxe.Verb, pbar.TOligos.origin, p.gxe.TOligos, labels = c("A", "", "B", "", "C", "", "D", ""), ncol = 2, align="hv")
```

## --  Total Dietary Fiber (TDF)

```{r fig.width = 8, fig.height=3, fig.align='center', echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Entry means by origin and lineplot of entry means by origin showing gxe of total dietary fiber. Bars with the same letter do not differe significantly between locations. An analysis of variance revealed there is not a significant effect of location (p=0.0589). There is a significant effect of gxe (p=0.015)."}
source('lineplot.gxe.R')
source('barplot.gxe.origin.R')
plot_grid(pbar.TDF.origin, p.gxe.TDF, ncol = 2, align="hv")
```









