---
output: 
  pdf_document:
    keep_tex: true
    fig_caption: true
    latex_engine: pdflatex
    template: ~/GitHub/manuscripts/bean_fiber/svm-latex-ms.tex
    toc: yes
    toc_depth: 2
title: "Variation of dietary fiber content in 282 common bean genotypes from the middle
  american gene pool"
thanks: ""
author:
- name: 
  affiliation: 
- name: 
  affiliation: 
- name: 
  affiliation: 
abstract: "Abstract content here..."
keywords: "common bean, fiber, oligosaccharide"
date: "`r format(Sys.time(), '%B %d, %Y')`"
geometry: margin=1in
fontfamily: mathpazo
mainfont: cmss
fontsize: 11pt
graphics: yes
# spacing: double
biblio-style: apsr
---
## Introduction

`r nrow(fiber)/2` entries from the BeanCAP study were grown in Fort Collins, CO in 2015 to determine variation in fiber content. All entries originated from the Middle American gene pool and were further subdivided into the following races and market classes: 

| Race | # Entries |
-------|----------
Durango | `r sum(fiber$race=="durango", na.rm=TRUE)/2`
Jalisco | `r sum(fiber$race=="jalisco", na.rm=TRUE)/2`
Mesoamerican | `r sum(fiber$race=="mesoamerican", na.rm=TRUE)/2`

| Market Class | # Entries |
------------ | ----------
pinto | `r sum(fiber$mktclass=="pinto", na.rm=TRUE)/2`
navy | `r sum(fiber$mktclass=="navy", na.rm=TRUE)/2`
black | `r sum(fiber$mktclass=="black", na.rm=TRUE)/2`
GN | `r sum(fiber$mktclass=="GN", na.rm=TRUE)/2`
small red | `r sum(fiber$mktclass=="small red", na.rm=TRUE)/2`
pink | `r sum(fiber$mktclass=="pink", na.rm=TRUE)/2`
carioca | `r sum(fiber$mktclass=="carioca", na.rm=TRUE)/2`
tan | `r sum(fiber$mktclass=="tan", na.rm=TRUE)/2`
black mottle | `r sum(fiber$mktclass=="black mottle", na.rm=TRUE)/2`
flor de mayo | `r sum(fiber$mktclass=="flor de mayo", na.rm=TRUE)/2`
red mottle | `r sum(fiber$mktclass=="red mottle", na.rm=TRUE)/2`

## Summary Statistics
Below are summary statistics for all dietary fiber components for all entries.
```{r echo=FALSE}
summary(fiber[,7:14])
```

## Results: Entries

Tables below show the five entries with the greatest and five entries with the lowest values of each fiber component averaged across the two replications including standard deviation (sd), standard error (se) and confindence intervals (ci).

##### Insoluble Dietary Fiber (IDF)
```{r echo=FALSE, message=FALSE, warning=FALSE}
library(data.table)
fiber<- read.csv("fiber.csv")
source('summarySE.R')
summ.IDF <-summarySE(fiber, measurevar="IDF", groupvars=c("Entry","mktclass","race"))
summ.IDF <- data.frame(summ.IDF)
#summ.f <- summ.F[fiber$ftype=="IDF", ]
f.IDF <- subset(summ.IDF, select=c(Entry, mktclass, race, IDF, sd, se, ci))
newdata <- f.IDF[order(-f.IDF$IDF), ]
newdata[,4:7] <- round(newdata[,4:7], 2)
DT <- data.table(newdata)
head(DT, 5)
tail(DT,5)
```
##### Soluble Dietary Fiber (SDF)
```{r echo=FALSE, message=FALSE, warning=FALSE}
library(data.table)
fiber<- read.csv("fiber.csv")
source('summarySE.R')
summ.SDF <-summarySE(fiber, measurevar="SDF", groupvars=c("Entry","mktclass","race"))
summ.SDF <- data.frame(summ.SDF)
#summ.f <- summ.F[fiber$ftype=="SDF", ]
f.SDF <- subset(summ.SDF, select=c(Entry, mktclass, race, SDF, sd, se, ci))
newdata <- f.SDF[order(-f.SDF$SDF), ]
newdata[,4:7] <- round(newdata[,4:7], 2)
DT <- data.table(newdata)
head(DT, 5)
tail(DT,5)
```
##### Raffinose (Raff)
```{r echo=FALSE, message=FALSE, warning=FALSE}
library(data.table)
fiber<- read.csv("fiber.csv")
source('summarySE.R')
summ.Raff <-summarySE(fiber, measurevar="Raff", groupvars=c("Entry","mktclass","race"))
summ.Raff <- data.frame(summ.Raff)
#summ.f <- summ.F[fiber$ftype=="Raff", ]
f.Raff <- subset(summ.Raff, select=c(Entry, mktclass, race, Raff, sd, se, ci))
newdata <- f.Raff[order(-f.Raff$Raff), ]
newdata[,4:7] <- round(newdata[,4:7], 2)
DT <- data.table(newdata)
head(DT, 5)
tail(DT,5)
```
##### Stachyose (Stach)
```{r echo=FALSE, message=FALSE, warning=FALSE}
library(data.table)
fiber<- read.csv("fiber.csv")
source('summarySE.R')
summ.stach <-summarySE(fiber, measurevar="Stach", groupvars=c("Entry","mktclass","race"))
summ.stach <- data.frame(summ.stach)
#summ.f <- summ.F[fiber$ftype=="Stach", ]
f.stach <- subset(summ.stach, select=c(Entry, mktclass, race, Stach, sd, se, ci))
newdata <- f.stach[order(-f.stach$Stach), ]
newdata[,4:7] <- round(newdata[,4:7], 2)
DT <- data.table(newdata)
head(DT, 5)
tail(DT,5)
```
##### Verbascose (Verb)
```{r echo=FALSE, message=FALSE, warning=FALSE}
library(data.table)
fiber<- read.csv("fiber.csv")
source('summarySE.R')
summ.Verb <-summarySE(fiber, measurevar="Verb", groupvars=c("Entry","mktclass","race"))
summ.Verb <- data.frame(summ.Verb)
#summ.f <- summ.F[fiber$ftype=="Verb", ]
f.Verb <- subset(summ.Verb, select=c(Entry, mktclass, race, Verb, sd, se, ci))
newdata <- f.Verb[order(-f.Verb$Verb), ]
newdata[,4:7] <- round(newdata[,4:7], 3)
DT <- data.table(newdata)
head(DT, 5)
tail(DT,5)
```
##### Total Oligosachharides (TOligos) = Raff + Stach + Verb
```{r echo=FALSE, message=FALSE, warning=FALSE}
library(data.table)
fiber<- read.csv("fiber.csv")
source('summarySE.R')
summ.TOligos <-summarySE(fiber, measurevar="TOligos", groupvars=c("Entry","mktclass","race"))
summ.TOligos <- data.frame(summ.TOligos)
#summ.f <- summ.F[fiber$ftype=="TOligos", ]
f.TOligos <- subset(summ.TOligos, select=c(Entry, mktclass, race, TOligos, sd, se, ci))
newdata <- f.TOligos[order(-f.TOligos$TOligos), ]
newdata[,4:7] <- round(newdata[,4:7], 2)
DT <- data.table(newdata)
head(DT, 5)
tail(DT,5)
```
##### Total Dietary Fiber (TDF) = IDF + SDF + Raff + Stach + Verb
```{r echo=FALSE, message=FALSE, warning=FALSE}
library(data.table)
fiber<- read.csv("fiber.csv")
source('summarySE.R')
summ.TDF <-summarySE(fiber, measurevar="TDF", groupvars=c("Entry","mktclass","race"))
summ.TDF <- data.frame(summ.TDF)
#summ.f <- summ.F[fiber$ftype=="TDF", ]
f.TDF <- subset(summ.TDF, select=c(Entry, mktclass, race, TDF, sd, se, ci))
newdata <- f.TDF[order(-f.TDF$TDF), ]
newdata[,4:7] <- round(newdata[,4:7], 2)
DT <- data.table(newdata)
head(DT, 5)
tail(DT,5)
```


## Results: Marketclass

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(data.table)
fiber.one<- read.csv("fiber_facet.csv")
```

| mktclass | IDF | 
---------|--------
black | `r mean(fiber.one$mktclass=="black" & fiber.one$ftype=="IDF", na.rm=TRUE)`a


##   -   Boxplots
Figures below depict fiber content by market class. 

```{r fig.width = 8, fig.height=8, fig.align='center', echo=TRUE, message=FALSE, warning=FALSE, fig.cap="Oligosaccharide content in 282 common bean genotypes by marketclass (A) Raffinose, (B) Stachyose, (C) Verbascose, and (D) total oligosaccharides (= Raff + Stach + Verb)."}
##Boxplots by Marketclass: Oligosaccharides
source('boxplot.mktclass.R')
plot_grid(p.Verb.mkt, p.Raff.mkt, p.Stach.mkt, p.TOligos.mkt, labels = c("A", "B", "C", "D"), ncol = 2)
```

```{r fig.width = 8, fig.height=5, fig.align='center', echo=FALSE, message=FALSE, warning=FALSE, fig.cap="(A) Soluble and (B) insoluble dietary fiber in 282 common bean genotypes by marketclass,"}
##Boxplots by Marketclass: Soluble and Insoluble DF
source('boxplot.mktclass.R')
plot_grid(p.IDF.mkt, p.SDF.mkt, labels = c("A", "B"), ncol = 2)
```

```{r fig.width = 4, fig.height=5, fig.align='center', echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Total Dietary Fiber content in 282 common bean genotypes by marketclass"}
##Boxplots by Marketclass: Total Dietary Fiber
source('boxplot.mktclass.R')
p.TDF.mkt
```



## Results: Race

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(data.table)
fiber.one<- read.csv("fiber_facet.csv")
f.idf <- fiber.one[which(fiber.one$ftype=="IDF") ,]
f.sdf <- fiber.one[which(fiber.one$ftype=="SDF") ,]
f.raff <- fiber.one[which(fiber.one$ftype=="Raff") ,]
f.stach <- fiber.one[which(fiber.one$ftype=="Stach") ,]
f.verb <- fiber.one[which(fiber.one$ftype=="Verb") ,]
f.toligos <- fiber.one[which(fiber.one$ftype=="TOligos") ,]
f.tdf <- fiber.one[which(fiber.one$ftype=="TDF") ,]
```

| Race | IDF | Verb | Raff | Stach | TOligos | TDF
---------|--------|--------|---------|---------|--------|----------
Durango | `r round(mean(f.idf$race=="durango", na.rm=TRUE),2)`a | `r round(mean(f.sdf$race=="durango", na.rm=TRUE),2)`a | `r round(mean(f.raff$race=="durango", na.rm=TRUE),2)`a | `r round(mean(f.stach$race=="durango", na.rm=TRUE),2)`a | `r round(mean(f.verb$race=="durango", na.rm=TRUE),2)`a | `r round(mean(f.toligos$race=="durango", na.rm=TRUE),2)`a | `r round(mean(f.tdf$race=="durango", na.rm=TRUE),2)`a |
Jalisco | `r round(mean(fiber.one$race=="jalisco" & fiber.one$ftype=="IDF", na.rm=TRUE),2)`b
Mesoamerican | `r round(mean(fiber.one$race=="mesoamerican" & fiber.one$ftype=="IDF", na.rm=TRUE),2)`a

##   -   Boxplots
Figures below depict fiber content by race.

```{r fig.width = 8, fig.height=10, fig.align='center', echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Oligosaccharide content in 282 common bean genotypes by race (A) Raffinose, (B) Stachyose, (C) Verbascose, and (D) total oligosaccharides (= Raff + Stach + Verb)."}
##Boxplots by race: Oligosaccharides
source('boxplot.race.R')
plot_grid(p.Verb.race, p.Raff.race, p.Stach.race, p.TOligos.race, labels = c("A", "B", "C", "D"), ncol = 2)
```

```{r fig.width = 8, fig.height=5, fig.align='center', echo=FALSE, message=FALSE, warning=FALSE, fig.cap="(A) Soluble and (B) insoluble dietary fiber in 282 common bean genotypes by race,"}
##Boxplots by race: Soluble and Insoluble DF
source('boxplot.race.R')
plot_grid(p.IDF.race, p.SDF.race, labels = c("A", "B"), ncol = 2)
```

```{r fig.width = 2, fig.height=3, fig.align='center', echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Total Dietary Fiber content in 282 common bean genotypes by race"}
##Boxplots by race: Total Dietary Fiber
source('boxplot.race.R')
p.TDF.race
```